---
title: "COViD by the Numbers"
author: "Robert Z. Selden, Jr."
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: covid.bib
fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Because making informed decisions matters

Those data used on this page were harvested from the [Texas Department of State Health Services](https://www.dshs.state.tx.us/coronavirus/additionaldata.aspx) daily, and visualizations were modeled in R. This exercise was generated as a means for me to better understand the data, and the risk that COVID poses to my family. Knowledge is power in this situation, and---when the mortality of those that I love is at risk---I would like to be as well-informed and knowledgeable as possible. 

The text and visualizations listed below were built for my personal use only. I am working through automating this analysis code, and it will change as I progress in that effort. Will also include links and references to Texas-specific resources and case studies as I come across them.

### Daily cases in Texas

```{r totaltexascases, out.width = "100%", dpi = 600, echo=TRUE}
# load libraries
library(tidyverse)

# set working directory
setwd(getwd())

# load data
data <- read.csv("dailycases.csv")

# define variables
# reformat date column to meet with ISO 8601 (yyyy-mm-dd)
date <- as.Date(data$date,"%m/%d/%y")
# daily cases 
dailyc <- data$dailycases
# daily fatalities
dailyf <- data$dailyfatalities

# barplot of daily cases
p <- ggplot(data, aes(x=date, y=dailyc)) +
  geom_bar(stat = "identity", fill = "blue") +
  theme(axis.text.x = element_blank()) +
  labs(x = "March 26 - July 21, 2020", y = "Daily Cases in Texas") 
p
```

### Cumulative cases in Nacogdoches County

Cumulative numbers differ from the daily numbers reported to the State of Texas. I have not yet been able to find a dataset that includes the daily numbers for Nacogdoches County, but will continue looking.

```{r countycases, out.width = "100%", dpi = 600, echo=TRUE}
# data wrangling
# transpose data and extract the last five characters from the date column

# load data
data2 <- read.csv("dailycounty2.csv")

# define variables
# reformat date column to meet with ISO 8601 (yyyy-mm-dd)
date <- as.Date(data2$date,"%m/%d/%y")
# cumulative cases for Nacogdoches County
dailyc <- data2$cumulative

# barplot of daily cases
p <- ggplot(data2, aes(x=date, y=dailyc)) +
  geom_bar(stat = "identity", fill = "blue") +
  theme(axis.text.x = element_blank()) +
  labs(x = "March 25 - July 21, 2020", y = "Cumulative Cases in Nacgodoches County")
p
# end of code
```

